### 2021-03-17 Manuscript.Rmd
Load data

---
title: "Storm Data Analysis - Week 4"
author: "Marco Baciarello"
date: "3/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)

suppressMessages(library(tidyverse))
```

## Abstract

Blah blah

## Introduction

We analyze this, that and something else...

## Methods

The following code was used for analysis

### Data preparation

First, we download the data file and load it into an R object. We're using a `tidyr`-centric approach, leveraging on the full suite of library found in package `tidyverse`.

```{r data-prep}

# Check if data dir exists -> create one
if (!file.exists("data")) {
    dir.create("./data")
}

# Get the data
if (!file.exists("data/stormdata.csv")) {
    download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", "data/stormdata.csv")
}

data <- read_csv("data/stormdata.csv")

```

### 2021-03-20 Manuscript.Rmd
First round: data prep and changes to economical variables

@@
 suppressMessages(library(tidyverse))
+# library(skimr)
 ```
@@
 
-```{r data-prep}
+```{r data-prep, message=FALSE, warning=FALSE, cache = TRUE}
 
@@
 # Get the data
-if (!file.exists("data/stormdata.csv")) {
-    download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", "data/stormdata.csv")
+if (!file.exists("data/stormdata.csv.bz2")) {
+    download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", "data/stormdata.csv.bz2")
 }
 
-data <- read_csv("data/stormdata.csv")
+# Read data in. fread seems way faster than other solutions.
+data <- data.table::fread("data/stormdata.csv.bz2", data.table = FALSE)
 
 ```
+
+We create 2 new table columns; one is a properly-formatted `POSIXct` column including date and time information from columns `BGN_DATE`and `BGN_TIME`.
+A second column will be created in order to reconstruct information on property damage: columns `PROPDMGEXP` specifies an "exponent" for `PROPDMG`, in the form of a mathematical suffix: K for thousands, M for millions, and so on...
+We will compute a column containing the actual amount as an integer or double.
+Same operation will be performed for `CROPDMG` and `CROPDMGEXP`.
+
+```{r column-formatting}
+
+# Formatted data
+fdata <- data %>%
+    mutate(
+        # Paste date and time For some reason, data includes a fixed time 00:00
+        # which we don't need, so we split and get the first fragment
+        temp_datetime = paste(strsplit(BGN_DATE, " ")[[1]][1], BGN_TIME),
+        # Convert to datetime; 
+        utc_datetime = strptime(
+            temp_datetime,
+            "%m/%d/%Y %H%M"),
+        # correct timezones (there are older/unsupported abbreviations)
+        c_timezone = recode(TIME_ZONE,
+                            CDT = "US/Central",
+                            CST = "US/Central",
+                            UNK = "UTC",
+                            MDT = "US/Mountain",
+                            AST = "US/Eastern",
+                            ADT = "US/Eastern",
+                            CSt = "US/Central",
+                            CSC = "UTC",
+                            SCT = "US/Central",
+                            ESY = "US/Eastern",
+                            SST = "US/Samoa",
+                            AKS = "US/Alaska",
+                            GST = "US/Central"
+                            ),
+        # # Convert each time to its own time zone
+        datetime = lubridate::force_tzs(utc_datetime, tzones = c_timezone)
+    )
+```
+
+`*EXP` values are actually multipliers for variables indicating damages. The full legend can be found [here](ane.st/HandleExponent6c0f3 "Legend for EXP values").
+
+```{r exponents}
+unique(data$PROPDMGEXP)
+unique(data$CROPDMGEXP)
+```
+
+We will transform them into the actual multiplier so we can compute actual damages.
+It is known that there are a few malformed values [*e.g.*: empty strings ("")]; they will be set to 0.
+According to the above analysis, malformed values are negligible both in numbers and absolute values.
+
+```{r exp-to-multipliers}
+fdata <- fdata %>%
+    mutate(
+        prop_mult = case_when(
+            PROPDMGEXP %in% c(0, 1, 2, 3, 4, 5, 6, 7, 8) ~ 10,
+            PROPDMGEXP %in% c('m', 'M') ~ 1000000,
+            PROPDMGEXP %in% c('k', 'K') ~ 1000,
+            PROPDMGEXP %in% c('b', 'B') ~ 1000000000,
+            PROPDMGEXP %in% c('h', 'H') ~ 100,
+            PROPDMGEXP == '+' ~ 1,
+            TRUE ~ 0
+        ),
+        crop_mult = case_when(
+            PROPDMGEXP %in% c(0, 1, 2, 3, 4, 5, 6, 7, 8) ~ 10,
+            CROPDMGEXP %in% c('m', 'M') ~ 1000000,
+            CROPDMGEXP %in% c('k', 'K') ~ 1000,
+            CROPDMGEXP %in% c('b', 'B') ~ 1000000000,
+            CROPDMGEXP %in% c('h', 'H') ~ 100,
+            CROPDMGEXP == '+' ~ 1,
+            TRUE ~ 0
+        ),
+        prop_dmg = PROPDMG * prop_mult,
+        crop_dmg = CROPDMG * crop_mult
+    )
+
+```
+
+We can get the number of invalid `EXP` values resulting in `0`-value multipliers using the following code:
+
+```{r invalid-exps}
+# Property damage
+inv_prop_exp <- filter(fdata, PROPDMG != 0, prop_mult == 0) %>%
+    select(PROPDMG, PROPDMGEXP, prop_mult, BGN_DATE) %>%
+    nrow()
+
+inv_crop_exp <- filter(fdata, CROPDMG != 0, crop_mult == 0) %>%
+    select(CROPDMG, CROPDMGEXP, crop_mult, BGN_DATE) %>%
+    nrow()
+```
+
+Using **inline code evaluation**, we can report `r inv_prop_exp` invalid property damage exponents, and `r inv_crop_exp` invalid crop damage exponents.

### 2021-03-21 Manuscript.Rmd
- Switch to dtplyr - Computed and selected necessary variables for analysis

--- 2021-03-17 Manuscript.Rmd
@@
 suppressMessages(library(tidyverse))
+suppressMessages(library(dtplyr))
+suppressMessages(library(data.table))
+# library(skimr)
 ```
@@
 
-### Data preparation
+## Data preparation
 
@@
 # Get the data
-if (!file.exists("data/stormdata.csv")) {
-    download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", "data/stormdata.csv")
+if (!file.exists("data/stormdata.csv.bz2")) {
+    download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", "data/stormdata.csv.bz2")
 }
 
-data <- read_csv("data/stormdata.csv")
+# Read data in. fread seems way faster than other solutions.
+rdata <- fread("data/stormdata.csv.bz2", data.table = FALSE)
+# Convert to "lazy data.table" which will accept dplyr functions
+data <- lazy_dt(rdata)
 
 ```
+
+### Time formatting
+
+We create 2 new table columns; one is a properly-formatted `POSIXct` column including date and time information from columns `BGN_DATE` and `BGN_TIME`. We will ignore end times of phenomena for now.
+
+```{r column-formatting}
+
+# TECH NOTE: We're using the `dtplyr` interface to use `tidyverse` grammar but
+# run data.table code, for increased performance on large tables.
+
+# Format data
+fdata <- data %>%
+    mutate(
+        # Paste date and time For some reason, data includes a fixed time 00:00
+        # which we don't need, so we split and get the first fragment
+        temp_datetime = paste(strsplit(BGN_DATE, " ")[[1]][1], BGN_TIME),
+        # Convert to datetime; 
+        utc_datetime = as.POSIXct(
+            temp_datetime,
+            format = "%m/%d/%Y %H%M"),
+        # correct timezones (there are older/unsupported abbreviations)
+        c_timezone = recode(TIME_ZONE,
+                            CDT = "US/Central",
+                            CST = "US/Central",
+                            UNK = "UTC",
+                            MDT = "US/Mountain",
+                            AST = "US/Eastern",
+                            ADT = "US/Eastern",
+                            CSt = "US/Central",
+                            CSC = "UTC",
+                            SCT = "US/Central",
+                            ESY = "US/Eastern",
+                            SST = "US/Samoa",
+                            AKS = "US/Alaska",
+                            GST = "US/Central"
+                            ),
+        # # Convert each time to its own time zone
+        datetime = lubridate::force_tzs(utc_datetime, tzones = c_timezone),
+        # Force POSIXct format, otherwise data won't be usable in data.table
+        # package functions
+        begin_time = as.POSIXct(datetime)
+    ) %>%
+    # Need to drop non-POSIXlt columns for data.table compatibility
+    select(-contains("datetime")) %>%
+    # End dtplyr "session," translate into data.table code and run
+    as_tibble()
+
+```
+
+### Property and agricultural damages
+
+A second column will be created in order to reconstruct information on property damage: columns `PROPDMGEXP` specifies an "exponent" for `PROPDMG`, in the form of a mathematical suffix: K for thousands, M for millions, and so on...
+We will compute a column containing the actual amount as an integer or double.
+Same operation will be performed for `CROPDMG` and `CROPDMGEXP`.
+
+`*EXP` values are actually multipliers for variables indicating damages. The full legend can be found [here](ane.st/HandleExponent6c0f3 "Legend for EXP values").
+
+```{r exponents}
+unique(data$PROPDMGEXP)
+unique(data$CROPDMGEXP)
+```
+
+We will transform them into the actual multiplier so we can compute actual damages.
+It is known that there are a few malformed values [*e.g.*: empty strings ("")]; they will be set to 0.
+According to the above analysis, malformed values are negligible both in numbers and absolute values.
+
+```{r exp-to-multipliers}
+
+# Cannot use dtplyr here because of error: Can't rename variables in this context.
+
+mdata <- fdata %>%
+    mutate(
+        prop_mult = case_when(
+            PROPDMGEXP %in% c(0, 1, 2, 3, 4, 5, 6, 7, 8) ~ 10,
+            PROPDMGEXP %in% c('m', 'M') ~ 1000000,
+            PROPDMGEXP %in% c('k', 'K') ~ 1000,
+            PROPDMGEXP %in% c('b', 'B') ~ 1000000000,
+            PROPDMGEXP %in% c('h', 'H') ~ 100,
+            PROPDMGEXP == '+' ~ 1,
+            # Vector must be double because of limitations in `case_when`, so we
+            # can't set a value to NA; use dummy code and convert later.
+            TRUE ~ -999999
+        ),
+        crop_mult = case_when(
+            PROPDMGEXP %in% c(0, 1, 2, 3, 4, 5, 6, 7, 8) ~ 10,
+            CROPDMGEXP %in% c('m', 'M') ~ 1000000,
+            CROPDMGEXP %in% c('k', 'K') ~ 1000,
+            CROPDMGEXP %in% c('b', 'B') ~ 1000000000,
+            CROPDMGEXP %in% c('h', 'H') ~ 100,
+            CROPDMGEXP == '+' ~ 1,
+            # Vector must be double because of limitations in `case_when`, so we
+            # can't set a value to NA; use dummy code and convert later.
+            TRUE ~ -999999
+        ),
+        # Change dummy value into NA
+        across(ends_with("_mult"), ~na_if(., -999999)),
+        crop_dmg = CROPDMG * crop_mult,
+        prop_dmg = PROPDMG * prop_mult,
+        tot_dmg = prop_dmg + crop_dmg
+    )
+
+```
+
+We can get the number of invalid `EXP` values resulting in `0`-value multipliers using the following code:
+
+```{r invalid-exps}
+# Property damage
+inv_prop_exp <- filter(mdata, PROPDMG != 0, prop_mult == 0) %>%
+    select(PROPDMG, PROPDMGEXP, prop_mult, BGN_DATE) %>%
+    nrow()
+
+inv_crop_exp <- filter(mdata, CROPDMG != 0, crop_mult == 0) %>%
+    select(CROPDMG, CROPDMGEXP, crop_mult, BGN_DATE) %>%
+    nrow()
+```
+
+Using **inline code evaluation**, we can report `r inv_prop_exp` invalid property damage exponents, and `r inv_crop_exp` invalid crop damage exponents.
+Data from these was set to NA in order to compute means *etc...*
+
+### Injuries and fatalities
+
+These data do not seem to be needing any significant adjustments:
+
+```{r inj-fat-describe}
+
+skimr::skim(fdata$INJURIES)
+cat("\n\n\n")
+skimr::skim(fdata$FATALITIES)
+
+```
+
+### Selection of variables of interest
+
+According to the requirements for this report, we will be describing associations between weather events and human and economical damages.
+After preparatory operations performed above, we will now drop variables which are not of interest to these analyses, in order to make data more easily readable and reduce resource requirements.
+
+```{r data-column-thinning}
+
+# Select useful columns
+lean_data <- mdata %>%
+    select(REFNUM, EVTYPE, begin_time, c_timezone,
+           FATALITIES, INJURIES, tot_dmg)
+# Economical damages aren't always available, so better have a tibble with all
+# complete cases
+dmg_data <- lean_data %>%
+    filter(!is.na(tot_dmg))
+
+```
+
+### Selection of events
+
+The vast majority of events have caused no damage to people or property.
+More than 90% of event *types* are not associated with damage, and many of them have occurred just once since 1950.
+This is mostly due to numerous, uncorrected mistakes made in data entry over the first ⅔ of the database's life.
+
+We will now plot the distribution of values of interests, so as to only select significant *event types* in terms of occurrence. We'll work with the 90^th^ percentile (and above) of event frequency.
+
+```{r evtype-selection}
+
+# Make a copy of data and cast as DT
+dt_data <- as.data.table(lean_data)
+
+sel_data <- dt_data[, .(count = .N,
+                        total_dmg = sum(tot_dmg, na.rm = TRUE),
+                        fatalities = sum(FATALITIES, na.rm = TRUE),
+                        injuries = sum(INJURIES, na.rm = TRUE)
+                        ),
+                    by = EVTYPE][
+                        order(-count, -total_dmg, -fatalities)]
+
+View(sel_data)
+
+```

### 2021-03-21 Manuscript.Rmd
- Completed filtering and ranking - Moved reusable code to function - Intent is to apply ranking to datasets, differing on year-range (primary driver of variation between event types)

@@
 # library(skimr)
+```
+
+```{r function-definitions}
+
+# WARNING: operates by reference, modifies argument data.table Adds columns for
+# ranks based on event count, damage and injuries/fatalities. Useful to quickly
+# re-rank tables with different elements or filters based on year.
+rank_DT <- function(DT) {
+    DT[, .(
+            counter = .N, # Count elements in group
+            # Total monetary damages per group
+            total_dmg = sum(tot_dmg, na.rm = TRUE),
+            # Total fatalities
+            fatalities = sum(FATALITIES, na.rm = TRUE),
+            # Injuries
+            injuries = sum(INJURIES, na.rm = TRUE)
+            ),
+        # Grouping by event type
+        by = event][
+            ,
+            # Compute a "percent rank": rank events by number of
+            # occurrences (desc), then assign a percentile by
+            # dividing rank by denominator (i.e. nrows)
+            # Note double dot syntax to access calling scope
+            `:=`(perc_rank = frankv(counter)/..denominator,
+                 # Same for monetary damage
+                 dmg_rank = frankv(total_dmg)/..denominator,
+                 # And life damage
+                 inj_rank = frankv(injuries)/..denominator,
+                 fat_rank = frankv(fatalities)/..denominator
+                 )
+            ][
+                # Filter for frequency rank >=90th
+                perc_rank >= 0.9
+            ][
+                order(-counter)
+            ]
+}
+
 ```
@@
         # package functions
-        begin_time = as.POSIXct(datetime)
+        begin_time = as.POSIXct(datetime),
+        # Finally, change EVTYPE into a factor ordered by frequency (may be
+        # CPU-intensive)
+        event = fct_infreq(as.factor(EVTYPE))
     ) %>%
@@
     mutate(
+        year = as.integer(lubridate::year(begin_time)),
         prop_mult = case_when(
@@
     ) 
-
 ```
@@
 lean_data <- mdata %>%
-    select(REFNUM, EVTYPE, begin_time, c_timezone,
+    select(REFNUM, event, year, c_timezone,
            FATALITIES, INJURIES, tot_dmg)
@@
 dt_data <- as.data.table(lean_data)
-
-sel_data <- dt_data[, .(count = .N,
+# Define the number of rows that will be in the grouped DT. This is equal to the
+# unique values of event, as we'll group by that.
+denominator <- uniqueN(dt_data, "event")
+# Apply final transformations to DT
+sel_data <- dt_data[, .(
+                        counter = .N, # Count elements in group
+                        # Total monetary damages per group
                         total_dmg = sum(tot_dmg, na.rm = TRUE),
+                        # Total fatalities
                         fatalities = sum(FATALITIES, na.rm = TRUE),
+                        # Injuries
                         injuries = sum(INJURIES, na.rm = TRUE)
                         ),
-                    by = EVTYPE][
-                        order(-count, -total_dmg, -fatalities)]
-
-View(sel_data)
+                    # Grouping by event type
+                    by = event][
+                        ,
+                        # Compute a "percent rank": rank events by number of
+                        # occurrences (desc), then assign a percentile by
+                        # dividing rank by denominator (i.e. nrows)
+                        # Note double dot syntax to access calling scope
+                        `:=`(perc_rank = frankv(counter)/..denominator,
+                             # Same for monetary damage
+                             dmg_rank = frankv(total_dmg)/..denominator,
+                             # And life damage
+                             inj_rank = frankv(injuries)/..denominator,
+                             fat_rank = frankv(fatalities)/..denominator
+                             )
+                        ][
+                            # Filter for frequency rank >=90th
+                            perc_rank >= 0.9
+                        ][
+                            order(-counter)
+                        ]
 

### 2021-03-22 Manuscript.Rmd
- Selected 95th percentile of overall data - Tests for panel plot

@@
             ]
+}
+
+# Min-max normalization
+normalize <- function(x) {
+    return ((x - min(x)) / (max(x) - min(x)))
   }
@@
                         ][
-                            # Filter for frequency rank >=90th
-                            perc_rank >= 0.9
+                            # Filter for frequency rank >=95th
+                            perc_rank >= 0.95
                         ][
@@
 ```
+
+Figure 1 is a panel plot of data at or above the 95^th^ percentile of (A) event frequency by type, (B) monetary damages and (C) damages to persons in the period of interest (1950–2013).
+One can easily that the magnitude of event-related damages decreases almost exponentially with frequency.
+It should be noted that the preponderant majority of events below 95^th^ percentile of frequency are actually data-entry errors arising from "event type" being an open-text field in the database.
+
+```{r frequency-plots}
+
+sel_data %>% ggplot(
+    aes(x = fct_reorder(event, injuries, .desc = T), y = inj_rank)
+    ) +
+    geom_point()
+
+```

### 2021-03-23 Manuscript.Rmd
Switched selection of cases to only include cases with at least one of: injuries, fatalities or damages

@@
 
-mdata <- fdata %>%
+temp_data <- fdata %>%
     mutate(
@@
     )
+
+# Switch back to dtplyr syntax and clean up
+temp_data <- lazy_dt(temp_data)
+
+# Filter out events which did not cause any injuries, fatalities or monetary
+# damages
+mdata <- temp_data %>% filter(
+  (INJURIES > 0) | 
+    (FATALITIES > 0) |
+    (tot_dmg > 0)
+) %>%
+  # Apply data.table code
+  as_tibble()
+
 ```
-
 We can get the number of invalid `EXP` values resulting in `0`-value multipliers using the following code:
@@
 # Property damage
-inv_prop_exp <- filter(mdata, PROPDMG != 0, prop_mult == 0) %>%
+inv_prop_exp <- filter(temp_data, PROPDMG != 0, prop_mult == 0) %>%
     select(PROPDMG, PROPDMGEXP, prop_mult, BGN_DATE) %>%
@@
 
-inv_crop_exp <- filter(mdata, CROPDMG != 0, crop_mult == 0) %>%
+inv_crop_exp <- filter(temp_data, CROPDMG != 0, crop_mult == 0) %>%
     select(CROPDMG, CROPDMGEXP, crop_mult, BGN_DATE) %>%
     nrow()
+
+# Clean up
+rm(temp_data)
 ```
@@
 
-skimr::skim(fdata$INJURIES)
+skimr::skim(mdata$INJURIES)
 cat("\n\n\n")
-skimr::skim(fdata$FATALITIES)
+skimr::skim(mdata$FATALITIES)
 
@@
                             # Filter for frequency rank >=95th
-                            perc_rank >= 0.95
+                            perc_rank >= 0.90
                         ][
@@
 sel_data %>% ggplot(
-    aes(x = fct_reorder(event, injuries, .desc = T), y = inj_rank)
+  aes(x = fct_reorder(event, perc_rank, .desc = T), y = counter)
 ) +
-    geom_point()
+  geom_point() +
+  scale_y_continuous(
+    # limits = c(0, 100), 
+    name = ("Frequency quantile\n")) +
+  theme_bw()
 

### 2021-03-23 Manuscript.Rmd
Completed element selection; started factor (event type) aggregation

@@
 # Min-max normalization
-normalize <- function(x) {
+normalizer <- function(x) {
     return ((x - min(x)) / (max(x) - min(x)))
@@
 ```{r exponents}
-unique(data$PROPDMGEXP)
-unique(data$CROPDMGEXP)
+unique(fdata$PROPDMGEXP)
+unique(fdata$CROPDMGEXP)
 ```
@@
 # Clean up
-rm(temp_data)
+# rm(temp_data)
 ```
@@
            FATALITIES, INJURIES, tot_dmg)
-# Economical damages aren't always available, so better have a tibble with all
-# complete cases
-dmg_data <- lean_data %>%
-    filter(!is.na(tot_dmg))
 
@@
 
-We will now plot the distribution of values of interests, so as to only select significant *event types* in terms of occurrence. We'll work with the 90^th^ percentile (and above) of event frequency.
-
 ```{r evtype-selection}
@@
 # Make a copy of data and cast as DT
-dt_data <- as.data.table(lean_data)
+sel_data <- as.data.table(lean_data)
 # Define the number of rows that will be in the grouped DT. This is equal to the
@@
 denominator <- uniqueN(dt_data, "event")
+# Make a copy for future use:
+tot_data <- copy(sel_data)
 # Apply final transformations to DT
-sel_data <- dt_data[, .(
+sel_data[, .(
     counter = .N, # Count elements in group
@@
 
-Figure 1 is a panel plot of data at or above the 95^th^ percentile of (A) event frequency by type, (B) monetary damages and (C) damages to persons in the period of interest (1950–2013).
-One can easily that the magnitude of event-related damages decreases almost exponentially with frequency.
-It should be noted that the preponderant majority of events below 95^th^ percentile of frequency are actually data-entry errors arising from "event type" being an open-text field in the database.
+```{r ev-selection-comparison}
 
-```{r frequency-plots}
+gdata <- as.data.table(temp_data)
+# With DT syntax, group and compute relevant sums
+totals <- gdata[,.(sum_inj = sum(INJURIES, na.rm = T),
+               sum_fat = sum(FATALITIES, na.rm = T),
+               sum_dmg = sum(tot_dmg, na.rm = T))]
 
-sel_data %>% ggplot(
-  aes(x = fct_reorder(event, perc_rank, .desc = T), y = counter)
-) +
-  geom_point() +
-  scale_y_continuous(
-    # limits = c(0, 100), 
-    name = ("Frequency quantile\n")) +
-  theme_bw()
+selected <- sel_data[,.(sum_inj = sum(INJURIES, na.rm = T),
+               sum_fat = sum(FATALITIES, na.rm = T),
+               sum_dmg = sum(tot_dmg, na.rm = T))]
+
+totals == selected
+```
+
+Events below the 90^th^ percentile of frequency are mostly irrelevant to the analysis.
+To confirm this, we calculated total damages, injuries and fatalities from the global data and from the selected data.
+Results are identical (see above), despite the latter table having only a fraction of the event types (`r length(unique(gdata$event))` _vs._ `r length(unique(sel_data$event))`).
+
+### Aggregation of events by type
+
+```{r event-aggregation}
+
+# Functions to aggregate (collapse) factor levels require specific values, specified in a named vector. Here, we use "data mining" to generate such vector.
+
+# Create regexp patterns to match to. The following is also the priority order
+# with which events will be assigned to a category.
+stormy <- c("storm*", "thunder*", "tstm", "rain", "precip")
+maritime <- c("sea*", "ocean*", "mari[nt]*", "curr*", "tid*", "surf*")
 

### 2021-03-23 Manuscript.Rmd
Trying to collapse event names into fewer ones

@@
         # CPU-intensive)
-        event = fct_infreq(as.factor(EVTYPE))
+        event = fct_infreq(as.factor(str_to_lower(EVTYPE)))
     ) %>%
@@
 # with which events will be assigned to a category.
-stormy <- c("storm*", "thunder*", "tstm", "rain", "precip")
-maritime <- c("sea*", "ocean*", "mari[nt]*", "curr*", "tid*", "surf*")
+new_fcts <- c(
+  hail = c("hail"),
+  tornado = c("torn", "nado"),
+  tsunami = c("tsun"),
+  stormy = c("storm*", "thunder*", "tstm", "rain", "precip", "hurric"),
+  lightning = c("lightn"),
+  flood = c("flood", "seiche"),
+  cold = c("cold", "ice", "sleet", "winter", "snow", "blizz", "avalan"),
+  heat = c("heat", "hot", "summ", "warm"),
+  windy = c("wind", "devil", "b*rst"),
+  maritime = c("sea*", "ocean*", "mari[nt]*",
+               "curr*", "tid*", "surf*", "wav", "swell"),
+  fog = c("fog"),
+  drought = c("drought"),
+  fire = c("fire", "blaze", "smok"),
+  landslide = c("slide")
+)
 
+# Initialize vector which will contain matching event names
+results <- c()
+counter <- 1
+
+for (gigio in new_fcts) {
+  fct <- str_subset(head(sel_data$event, 50), gigio)
+  results <- c(results, setNames(fct, names(new_fcts)[counter]))
+  counter <- counter + 1
+}
+
+names(results) <- res_names
 ```

### 2021-03-24 Manuscript.Rmd
Completed aggregation of event types

@@
 # Property damage
-inv_prop_exp <- filter(temp_data, PROPDMG != 0, prop_mult == 0) %>%
+inv_prop_exp <- filter(temp_data, PROPDMG != 0, prop_mult == -999999) %>%
   select(PROPDMG, PROPDMGEXP, prop_mult, BGN_DATE) %>%
+  as.data.frame() %>%
   nrow()
 
-inv_crop_exp <- filter(temp_data, CROPDMG != 0, crop_mult == 0) %>%
+inv_crop_exp <- filter(temp_data, CROPDMG != 0, crop_mult == -999999) %>%
   select(CROPDMG, CROPDMGEXP, crop_mult, BGN_DATE) %>%
+  as.data.frame() %>%
   nrow()
@@
 # unique values of event, as we'll group by that.
-denominator <- uniqueN(dt_data, "event")
+denominator <- uniqueN(fdata, "event")
 # Make a copy for future use:
@@
 
+We then aggregate event types into arbitrary categories, both for clarity and because there is artificial variability in the form of spelling errors and variations (_e.g._: )
+
 ```{r event-aggregation}
@@
 # with which events will be assigned to a category.
-new_fcts <- c(
+new_fcts <- list(
   hail = c("hail"),
@@
   tsunami = c("tsun"),
-  stormy = c("storm*", "thunder*", "tstm", "rain", "precip", "hurric"),
+  stormy = c("storm*", "thunder*", "tstm", "rain", "precip", "hurric", "spout"),
   lightning = c("lightn"),
@@
 # Initialize vector which will contain matching event names
-results <- c()
-counter <- 1
+old_events <- vector("list", 14)
+new_events <- c()
+i <- 1
 
-for (gigio in new_fcts) {
-  fct <- str_subset(head(sel_data$event, 50), gigio)
-  results <- c(results, setNames(fct, names(new_fcts)[counter]))
-  counter <- counter + 1
+suppressWarnings(
+  # For every new event category
+  for (i in seq(1:length(new_fcts))) {
+    # Match old event instances
+    old_instances <- str_subset(sel_data$event, new_fcts[[i]])
+    # Add to list
+    old_events[[i]] <- unique(old_instances)
+    i <- i + 1
   }
+)
+# Name (tag) list elements
+names(old_events) <- names(new_fcts)
 
-names(results) <- res_names
+# TODO: switch to using list names instead of manual entry of new factor levels
+sel_data$event2 <- fct_collapse(
+  sel_data$event,
+  hail = old_events$hail,
+  tornado = old_events$tornado,
+  tsunami = old_events$tsunami,
+  stormy = old_events$stormy,
+  lightning = old_events$lightning,
+  flood = old_events$flood,
+  cold = old_events$cold,
+  heat = old_events$heat,
+  windy = old_events$windy,
+  maritime = old_events$maritime,
+  fog = old_events$fog,
+  drought = old_events$drought,
+  fire = old_events$fire,
+  landslide = old_events$landslide,
+  other_level = "Other events"
+) %>%
+  fct_recode(
+    "Hail" = "hail",
+    "Winds" = "windy",
+    "Tornadoes" = "tornado",
+    "Floods" = "flood",
+    "Lightnings" = "lightning",
+    "Cold weather" = "cold",
+    "Storms/Rains" = "stormy",
+    "Oceanic events" = "maritime",
+    "Fires" = "fire",
+    "Droughts" = "drought",
+    "Hot weather" = "heat",
+    "Fog" = "fog",
+    "Landslides" = "landslide",
+    "Tsunamis" = "tsunami"
+  ) %>%
+  # Reorder according to frequency, desc
+  fct_infreq()
+
 ```

### 2021-03-24 Manuscript.Rmd
Messing with sel_data assignment by reference

@@
 
-## Methods
-
-The following code was used for analysis
-
-## Data preparation
+## Data Processing
 
@@
 A second column will be created in order to reconstruct information on property damage: columns `PROPDMGEXP` specifies an "exponent" for `PROPDMG`, in the form of a mathematical suffix: K for thousands, M for millions, and so on...
+Values can also be integers, in which case they seem to indicate that the reported damage value is to be multiplied by 10 (see full legend [here](ane.st/HandleExponent6c0f3 "Legend for EXP values").
 We will compute a column containing the actual amount as an integer or double.
 Same operation will be performed for `CROPDMG` and `CROPDMGEXP`.
-
-`*EXP` values are actually multipliers for variables indicating damages. The full legend can be found [here](ane.st/HandleExponent6c0f3 "Legend for EXP values").
 
@@
 ```
-We can get the number of invalid `EXP` values resulting in `0`-value multipliers using the following code:
 
-```{r invalid-exps}
+```{r invalid-exps, eval=FALSE, include=FALSE}
+
+# DEBUG: this chunk not to be run nor shown when knitting.
+
 # Property damage
@@
 ```
-
-Using **inline code evaluation**, we can report `r inv_prop_exp` invalid property damage exponents, and `r inv_crop_exp` invalid crop damage exponents.
-Data from these was set to NA in order to compute means *etc...*
 
@@
 To confirm this, we calculated total damages, injuries and fatalities from the global data and from the selected data.
-Results are identical (see above), despite the latter table having only a fraction of the event types (`r length(unique(gdata$event))` _vs._ `r length(unique(sel_data$event))`).
+Results are identical (see above), despite the latter table having only a fraction of the event types (`r length(unique(gdata$event))` *vs.* `r length(unique(sel_data$event))`).
 
@@
 
-We then aggregate event types into arbitrary categories, both for clarity and because there is artificial variability in the form of spelling errors and variations (_e.g._: )
+We then aggregate event types into arbitrary categories, both for clarity and because there is artificial variability in the form of spelling errors and variations (*e.g.*: `hail`, `marine hail`, `hail/wind`, `small hail` and so forth...)
 
@@
 
-# Initialize vector which will contain matching event names
+# Initialize list, vector and counter for the loop
 old_events <- vector("list", 14)
@@
 ```
+
+The choice of the fourteen categories was arbitrary, based on observation of event types appearing in the selected data.
+There are limitations to this approach, namely: events belonging to more than one category; and likely oversight of extremely rare events.
+
+```{r memory-release-cleanup}
+
+# Clean up large dataframes, no longer used
+erase_us <- c("data", "fdata", "gdata", "mdata", "rdata", "tot_data")
+rm(list = erase_us)
+```
+
+## Results
+
+```{r overall-plot}
+
+# Create plot data, use dtplyr for speed and convenience
+pdata <- lazy_dt(sel_data)
+
+# Plotting total instances of each event category
+# pdata %>%
+#     group_by(event2) %>%
+#     summarise(
+#         count = n(),
+#         
+#     )
+
+```

### 2021-03-24 Manuscript.Rmd
Changed event selection logic (new branch)

@@
 sel_data <- as.data.table(lean_data)
-# Define the number of rows that will be in the grouped DT. This is equal to the
-# unique values of event, as we'll group by that.
-denominator <- uniqueN(fdata, "event")
-# Make a copy for future use:
-tot_data <- copy(sel_data)
 # Apply final transformations to DT
-sel_data[, .(
+sel_data2 <- sel_data[, .(
     counter = .N, # Count elements in group
@@
   # Grouping by event type
-  by = event][
+  by = .(event)][
     ,
@@
     # Note double dot syntax to access calling scope
-    `:=`(perc_rank = frankv(counter)/..denominator,
+    `:=`(perc_rank = frankv(counter)/.N,
          # Same for monetary damage
@@
     # Filter for frequency rank >=95th
-    perc_rank >= 0.90
+    perc_rank >= 0.80
   ][
@@
 
-selected <- sel_data[,.(sum_inj = sum(INJURIES, na.rm = T),
-               sum_fat = sum(FATALITIES, na.rm = T),
-               sum_dmg = sum(tot_dmg, na.rm = T))]
+selected <- sel_data2[,.(sum_inj = sum(injuries, na.rm = T),
+               sum_fat = sum(fatalities, na.rm = T),
+               sum_dmg = sum(total_dmg, na.rm = T))]
 
-totals == selected
+# Proportion of damages in selected events as opposed to total
+round((selected/totals), digits = 2)
 ```
 
-Events below the 90^th^ percentile of frequency are mostly irrelevant to the analysis.
+Events below the 80^th^ percentile of frequency are mostly irrelevant to the analysis.
 To confirm this, we calculated total damages, injuries and fatalities from the global data and from the selected data.
-Results are identical (see above), despite the latter table having only a fraction of the event types (`r length(unique(gdata$event))` *vs.* `r length(unique(sel_data$event))`).
+Results are identical (see above), despite the latter table having only a fraction of the event types (`r length(unique(gdata$event))` *vs.* `r length(unique(sel_data2$event))`).
 
@@
 # Clean up large dataframes, no longer used
-erase_us <- c("data", "fdata", "gdata", "mdata", "rdata", "tot_data")
+erase_us <- c("data", "fdata", "gdata", "mdata", "rdata")
 rm(list = erase_us)

### 2021-03-24 Manuscript.Rmd
Completed event aggregation

@@
          # Same for monetary damage
-         dmg_rank = frankv(total_dmg)/..denominator,
+         dmg_rank = frankv(total_dmg)/.N,
          # And life damage
-         inj_rank = frankv(injuries)/..denominator,
-         fat_rank = frankv(fatalities)/..denominator
+         inj_rank = frankv(injuries)/.N,
+         fat_rank = frankv(fatalities)/.N
     )
@@
 To confirm this, we calculated total damages, injuries and fatalities from the global data and from the selected data.
-Results are identical (see above), despite the latter table having only a fraction of the event types (`r length(unique(gdata$event))` *vs.* `r length(unique(sel_data2$event))`).
+The ratios of the selected data to the global data is near 1 (see above), despite the selected data table having only a fraction of the event types (`r length(unique(gdata$event))` *vs.* `r length(unique(sel_data2$event))`).
 
@@
 
-```{r event-aggregation}
+```{r event-aggregation, warning=FALSE}
 
@@
   lightning = c("lightn"),
-  flood = c("flood", "seiche"),
-  cold = c("cold", "ice", "sleet", "winter", "snow", "blizz", "avalan"),
+  flood = c("flood", "seiche", "fld"),
+  cold = c("snow", "cold", "ice", "sleet", "^*inter", "blizz", "avalan"),
   heat = c("heat", "hot", "summ", "warm"),
   windy = c("wind", "devil", "b*rst"),
+  fog = c("fog"),
   maritime = c("sea*", "ocean*", "mari[nt]*",
                "curr*", "tid*", "surf*", "wav", "swell"),
-  fog = c("fog"),
   drought = c("drought"),
@@
     for (i in seq(1:length(new_fcts))) {
+        # Prepare storage array
+        old_instances <- c()
+        k <- 1
+        # Loop over elements in list element
+        for (k in seq(1:length(new_fcts[[i]]))) {
             # Match old event instances
-    old_instances <- str_subset(sel_data$event, new_fcts[[i]])
+            old_instances <- c(old_instances,
+                               str_subset(sel_data2$event, new_fcts[[i]][k])
+                                )
+            k <- k + 1
+        }
         # Add to list
@@
 # TODO: switch to using list names instead of manual entry of new factor levels
-sel_data$event2 <- fct_collapse(
-  sel_data$event,
+sel_data2$event2 <- fct_collapse(
+  sel_data2$event,
   hail = old_events$hail,
@@
 # Create plot data, use dtplyr for speed and convenience
-pdata <- lazy_dt(sel_data)
+pdata <- lazy_dt(sel_data2)
 

### 2021-03-25 Manuscript.Rmd
Final plot (total dmg)

@@
 The choice of the fourteen categories was arbitrary, based on observation of event types appearing in the selected data.
-There are limitations to this approach, namely: events belonging to more than one category; and likely oversight of extremely rare events.
+There are limitations to this approach, namely: events belonging to more than one category; and possible oversight of extremely rare events.
 
@@
 
-# Create plot data, use dtplyr for speed and convenience
-pdata <- lazy_dt(sel_data2)
+# Make a copy of data to be sorted by damage total, descending, for use in
+# damage-related plots.
+dmgdata <- sel_data2 %>%
+    group_by(event2) %>%
+    summarise(total_dmg = sum(total_dmg)) %>%
+    mutate(event2 = as.character(event2)) %>%
+    arrange(desc(total_dmg)) %>%
+    as_tibble()
+
+dmgdata %>%
+    ggplot(aes(
+        x = reorder(event2, -total_dmg),
+        y = total_dmg)) +
+    geom_col() +
+    scale_y_continuous(
+        labels = scales::label_dollar(prefix = NULL,
+                                      scale = 1e-9,
+                                      largest_with_cents = 1),
+        name = "Total estimated damages caused\n(in billions of dollars)\n"
+    ) + 
+    scale_x_discrete("\nWeather event category (1950–2013)") +
+    coord_cartesian(ylim = c(0, 5e10)) +
+    theme_classic()
 

### 2021-03-28 Manuscript.Rmd
Finalization

--- 2021-03-24 Manuscript.Rmd
@@
 output: html_document
+  fig_caption: yes
 ---
@@
 
+Nested loops is our way of going through a factor variable and associating each value to a value in a new factor variable (many-to-one relationship.).
+The need for loops seems to be related to the fact that these vars are stored as factors, and the lack of a specific function in the forcats package; fct_collapse() might have been one solution, but it takes the exact factor level names—since there were almost 1000 different possible levels, it would have been impractical to call this function.
+
 ```{r event-aggregation, warning=FALSE}
@@
 i <- 1
+```
 
+We create a list of possible synonims for weather events.
+In the inner loop, the factor variable is run (using `stringr::subset()` sequentially against each list element and its children.
+Where there's a match, the function returns the full value of the variable.
+We aggregate these values in a vector.
+In the outer loop, the unique values (i.e., factor levels) are appended to a second list.
+The second list has a list-element for each of the new factor levels; the old factor levels are thus stored as children of the new ones, based on matching with our criteria.
+
+Once complete, the list _could_ be used as a named vector. However, for some reason we had to insert each element manually using list$element syntax.
+
+```{r factor-loops}
 suppressWarnings(
@@
 ) %>%
+  # Adjust spelling/capitalization of factor levels
   fct_recode(
@@
 
+```{r figure-captions}
+fig.1.caption = "**Fig. 1:** Figure 1"
+```
+
 ## Results
 
-```{r overall-plot}
+Data on agricultural (crop) and property damages show unequivocal predominance of flooding phenomena as causes.
+At `r scales::dollar(dmgdata$total_dmg[1])`, overflowing of oceans, lakes and rivers caused about 4 times as much damage as storms and other rain phenomena, a distant second in ranking.
+One can easily imagine damages from disasters such as hurricanes Katrina and Sandy are the primary drivers of this result.
 
-# Create plot data, use dtplyr for speed and convenience
-pdata <- lazy_dt(sel_data2)
+```{r damage-plot, fig.cap=fig.1.caption}
 
-# Plotting total instances of each event category
-# pdata %>%
-#     group_by(event2) %>%
-#     summarise(
-#         count = n(),
-#         
-#     )
+# Make a copy of data to be sorted by damage total, descending, for use in
+# damage-related plots.
+dmgdata <- sel_data2 %>%
+    group_by(event2) %>%
+    summarise(total_dmg = sum(total_dmg),
+              total_inj = sum(injuries),
+              total_fat = sum(fatalities)
+              ) %>%
+    mutate(event2 = as.character(event2)) %>%
+    arrange(desc(total_dmg)) %>%
+    as_tibble()
+
+### Damage plot
+
+dmgdata %>%
+  ggplot(aes(
+    x = reorder(event2, -total_dmg),
+    y = total_dmg)) +
+  geom_col() +
+  scale_y_continuous(
+    labels = scales::label_dollar(prefix = NULL,
+                                  scale = 1e-9,
+                                  largest_with_cents = 1),
+    name = "Total estimated damages caused\n(in billions of dollars)\n"
+  ) + 
+  scale_x_discrete("\nWeather event category (1950–2013)") +
+  coord_cartesian(ylim = c(0, 5e10)) +
+  annotate("segment", x= 2.2, xend = 1.5, y = 50e9, yend = 52e9,
+           arrow = arrow(angle = 20,
+                         length = unit(0.15, "inches"),
+                         type = "closed")
+  ) + 
+  annotate("text",
+           x = 4.3, y = 49e9,
+           label = "Plot clipped for readability.\nActual value: $147 billion") +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 70, hjust = 1))
+#####
 
 ```
+
+```{r injuries-plot}
+### Injuries plot
+
+dmgdata %>%
+    ggplot(aes(
+        x = reorder(event2, -total_inj),
+        y = total_inj)) +
+    geom_col() +
+    scale_x_discrete("\nWeather event category (1950–2013)") +
+    scale_y_continuous(name = "Related injuries reported\n",
+                       breaks = seq(0, 1e5, 1e4),
+                       labels = scales::label_comma()) +
+    theme_classic() +
+    theme(axis.text.x = element_text(angle = 70, hjust = 1))
+
+```
+
+```{r fatalities-plot}
+### Fatalities plot
+
+dmgdata %>%
+    ggplot(aes(
+        x = reorder(event2, -total_fat),
+        y = total_fat)) +
+    geom_col() +
+    scale_x_discrete("\nWeather event category (1950–2013)") +
+    scale_y_continuous(name = "Related fatalities reported\n",
+                       #breaks = scales::extended_breaks(),
+                       breaks = seq(0, 5e3, 1e3),
+                       labels = scales::label_comma()) +
+    theme_classic() +
+    theme(axis.text.x = element_text(angle = 70, hjust = 1))
+```
+
+## Conclusions

### 2021-03-28 Manuscript.Rmd
Final, start bug checks

@@
 date: "3/17/2021"
-output: html_document
-  fig_caption: yes
+output:
+  html_document:
+    fig_captions: TRUE
+
 ---
@@
 
-```{r figure-captions}
+```{r figure-captions, include=FALSE}
 fig.1.caption = "**Fig. 1:** Figure 1"
@@
 
-## Results
-
-Data on agricultural (crop) and property damages show unequivocal predominance of flooding phenomena as causes.
-At `r scales::dollar(dmgdata$total_dmg[1])`, overflowing of oceans, lakes and rivers caused about 4 times as much damage as storms and other rain phenomena, a distant second in ranking.
-One can easily imagine damages from disasters such as hurricanes Katrina and Sandy are the primary drivers of this result.
-
-```{r damage-plot, fig.cap=fig.1.caption}
+``` {r final-data-computations}
 
@@
     as_tibble()
+
+```
+
+## Results
+
+Data on agricultural (crop) and property damages show unequivocal predominance of flooding phenomena as causes.
+At `r scales::dollar(dmgdata$total_dmg[1])`, overflowing of oceans, lakes and rivers caused about 4 times as much damage as storms and other rain phenomena, a distant second in ranking.
+One can easily imagine damages from disasters such as hurricanes Katrina and Sandy are the primary drivers of this result.
+
+```{r damage-plot, fig.cap=fig.1.caption}
 

### 2021-03-28 Manuscript.Rmd
Final, start bug checks

@@
 ---
-title: "Storm Data Analysis - Week 4"
+title: "Weather-Related Damages to Life and Property in the United States: a Retrospective Analysis from National Weather Service Data Between 1950 and 2011"
 author: "Marco Baciarello"
-date: "3/17/2021"
+date: "3/28/2021"
 output:
@@
 
-```{r function-definitions}
+## Synopsis
 
-# WARNING: operates by reference, modifies argument data.table Adds columns for
-# ranks based on event count, damage and injuries/fatalities. Useful to quickly
-# re-rank tables with different elements or filters based on year.
-rank_DT <- function(DT) {
-    DT[, .(
-            counter = .N, # Count elements in group
-            # Total monetary damages per group
-            total_dmg = sum(tot_dmg, na.rm = TRUE),
-            # Total fatalities
-            fatalities = sum(FATALITIES, na.rm = TRUE),
-            # Injuries
-            injuries = sum(INJURIES, na.rm = TRUE)
-            ),
-        # Grouping by event type
-        by = event][
-            ,
-            # Compute a "percent rank": rank events by number of
-            # occurrences (desc), then assign a percentile by
-            # dividing rank by denominator (i.e. nrows)
-            # Note double dot syntax to access calling scope
-            `:=`(perc_rank = frankv(counter)/..denominator,
-                 # Same for monetary damage
-                 dmg_rank = frankv(total_dmg)/..denominator,
-                 # And life damage
-                 inj_rank = frankv(injuries)/..denominator,
-                 fat_rank = frankv(fatalities)/..denominator
-                 )
-            ][
-                # Filter for frequency rank >=90th
-                perc_rank >= 0.9
-            ][
-                order(-counter)
-            ]
-}
-
-# Min-max normalization
-normalizer <- function(x) {
-    return ((x - min(x)) / (max(x) - min(x)))
-  }
-
-```
-
-## Abstract
-
-Blah blah
+In this data analysis we studied data from the National Weather Service, detailing weather events and the damages they caused between 1950 and 2011.
+Most data entry was poorly regulated, especially during the database's early years, so text mining was required in order to classify weather events in more generalizable categories.
+From this database, it appears that floods were by far the main cause of monetary losses in the studied period; it is likely that Hurricane Katrina and, to a lesser extent, Sandy were the main drivers of this result.
+Injuries and loss of life occurred significantly more often as consequences of tornadoes.
+Other wind events, extreme temperatures and floods were also associated with significant loss of life.
 
@@
 
-```{r figure-captions, include=FALSE}
-fig.1.caption = "**Fig. 1:** Figure 1"
-```
-
 ``` {r final-data-computations}
@@
 
+```{r figure-captions, include=FALSE}
+fig.1.caption = "**Fig.
+1:** Total damages (from lost crops and property) by weather event category.
+For readability, the first column has been clipped instead of being changed to a logarithmic scale."
+
+fig.2.caption = "**Fig. 2:** Total injuries by weather event category."
+
+fig.3.caption = "**Fig. 3:** Total fatalities by weather event category."
+```
+
 ## Results
@@
 One can easily imagine damages from disasters such as hurricanes Katrina and Sandy are the primary drivers of this result.
+In fact, this category was so much higher that we chose to clip the plot in order to keep data from other categories readable—a logarithmic y-axis scale would also have been appropriate, but it would have been more difficult to interpret at first sight.
 
@@
   ) + 
-  scale_x_discrete("\nWeather event category (1950–2013)") +
+  scale_x_discrete("\nWeather event category (1950–2011)") +
   coord_cartesian(ylim = c(0, 5e10)) +
@@
   theme(axis.text.x = element_text(angle = 70, hjust = 1))
-#####
-
 ```
 
-```{r injuries-plot}
+Data about injuries and fatalities is significantly different.
+The main culprits here seem to be tornadoes, which are presumably associated with danger from flying object.
+Other types of wind events are between 2^nd^ and 4^th^ ranks.
+Extremely hot weather seems to have been an important cause of death (2^nd^ rank, with 50% as many victims as tornadoes).
+
+```{r injuries-plot, fig.cap=fig.2.caption}
 ### Injuries plot
@@
     geom_col() +
-    scale_x_discrete("\nWeather event category (1950–2013)") +
+    scale_x_discrete("\nWeather event category (1950–2011)") +
     scale_y_continuous(name = "Related injuries reported\n",
@@
 
-```{r fatalities-plot}
+```{r fatalities-plot, fig.cap=fig.3.caption}
 ### Fatalities plot
@@
     geom_col() +
-    scale_x_discrete("\nWeather event category (1950–2013)") +
+    scale_x_discrete("\nWeather event category (1950–2011)") +
     scale_y_continuous(name = "Related fatalities reported\n",
@@
 ## Conclusions
+
+Massive flooding has historically been associated with the greatest crop and property damages, with rain/storms a very distant second.
+Tornadoes are the most dangerous type of events in terms of human life.

### 2021-03-28 Manuscript.Rmd

@@
 
-```{r setup, include=FALSE}
+```{r setup}
 knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)
@@
 suppressMessages(library(data.table))
-# library(skimr)
 ```
@@
 Other wind events, extreme temperatures and floods were also associated with significant loss of life.
-
-## Introduction
-
-We analyze this, that and something else...
 
@@
   nrow()
-
-# Clean up
-# rm(temp_data)
 ```
@@
 ```{r inj-fat-describe}
-
 skimr::skim(mdata$INJURIES)
-cat("\n\n\n")
 skimr::skim(mdata$FATALITIES)
-
 ```
@@
   ]
-
 ```
@@
 
-```{r factor-loops}
+```{r factor-loops, message=FALSE, warning=FALSE}
 suppressWarnings(
@@
   fct_infreq()
-
 ```
@@
 
-```{r memory-release-cleanup}
+```{r memory-release-cleanup, include=FALSE}
 
